Introduction

### 2015-07-21 dissertation/problems.tex
Problems addressed

\begin{problem}
SPARQL has some highly complex parts.
\end{problem}

@@

\begin{problem}
Public endpoints may have unpredictable usage patterns.
\end{problem}

@@

\begin{problem}
The distribution of data is often dominated by a power-law.
\end{problem}
However, 
\begin{problem}
The data distribution is often unknown to the query engine, and so,
following the principle of indifference, assumes uniform
distributions.
\end{problem}

@@

\begin{problem}
SPARQL endpoints, which are based on HTTP, will only enable caching
syntactically identical queries by  default.
\end{problem}

\begin{problem}
Determining that two queries will return the exact same answer is a
complex problem.
\end{problem}

\begin{problem}
It is not known how much of the Semantic Web exposes information that
could be used in HTTP caching.
\end{problem}

@@

\begin{problem}
How can suitable inference engines be made available to all users?
\end{problem}

@@

\begin{problem}
With a centralised ontology approach, how are anybody enabled to say
anything?
\end{problem}

@@

\begin{problem}
How can the adoption of existing vocabularies be encouraged? 
% TODO: Find the ref by Frank on the economics
\end{problem}

@@

\begin{problem}
Evaluation methodologies are not founded on sound statistics.
\end{problem}

@@

\begin{problem}
Establish laws of the Semantic Web information universe can reasonably be
expected to have validity far beyond the present.
\end{problem}

@@

\begin{problem}
Evaluation methodologies have little foundation in philosophy of science.
\end{problem}

@@

\begin{problem}
By breaking down the query down to individual triple patterns, the
query engine cannot take advantage of optimisations that involve
multiple triple patterns or other parts of the query.
\end{problem}

\begin{problem}
If methods such as \pcode{get\_pattern} are added for every part of the
query, it would increase the complexity of the API dramatically.
\end{problem}

\begin{problem}
By merely passing the entire query representation to an
implementation, the burden of evaluating the entire query is also
transferred, which makes it harder to use default implementations for
most of the query.
\end{problem}

@@

\begin{problem}
Current programming paradigm deals poorly with dynamic RDF data
containing a mix of terminological and assertional information.
\end{problem}

>>>>

\begin{problem}\label{prob:sparqlcomplex}
SPARQL has some highly complex parts.
\end{problem}

@@

\begin{problem}\label{prob:endpointunpred}
Public endpoints may have unpredictable usage patterns.
\end{problem}

@@

\begin{problem}\label{prob:zipf}
The distribution of data is often dominated by a power-law.
\end{problem}
However, 
\begin{problem}\label{prob:uniform}
The data distribution is often unknown to the query engine, and so,
following the principle of indifference, assumes uniform
distributions.
\end{problem}

@@

\begin{problem}\label{prob:syntacticcache}
SPARQL endpoints, which are based on HTTP, will only enable caching
syntactically identical queries by  default.
\end{problem}

\begin{problem}\label{prob:queryeq}
Determining that two queries will return the exact same answer is a
complex problem.
\end{problem}

\begin{problem}\label{prob:sanity}
It is not known how much of the Semantic Web exposes information that
could be used in HTTP caching.
\end{problem}

@@

\begin{problem}\label{prob:infgivecode}
How can suitable inference engines be made available to all users?
\end{problem}

@@

\begin{problem}\label{prob:sayanything}
With a centralised ontology approach, how are anybody enabled to say
anything?
\end{problem}

@@

\begin{problem}\label{prob:useexist}
How can the adoption of existing vocabularies be encouraged? 
% TODO: Find the ref by Frank on the economics
\end{problem}

@@

\begin{problem}\label{prob:badstats}
Evaluation methodologies are not founded on sound statistics.
\end{problem}

@@

\begin{problem}\label{prob:lawsfuture}
Establish laws of the Semantic Web information universe can reasonably be
expected to have validity far beyond the present.
\end{problem}

@@

\begin{problem}\label{prob:badphil}
Evaluation methodologies have little foundation in philosophy of science.
\end{problem}

@@

\begin{problem}\label{prob:breakdown}
By breaking down the query down to individual triple patterns, the
query engine cannot take advantage of optimisations that involve
multiple triple patterns or other parts of the query.
\end{problem}

\begin{problem}\label{prob:complexapi}
If methods such as \pcode{get\_pattern} are added for every part of the
query, it would increase the complexity of the API dramatically.
\end{problem}

\begin{problem}\label{prob:dontjustpass}
By merely passing the entire query representation to an
implementation, the burden of evaluating the entire query is also
transferred, which makes it harder to use default implementations for
most of the query.
\end{problem}

@@

\begin{problem}\label{prob:dynaprog}
Current programming paradigm deals poorly with dynamic RDF data
containing a mix of terminological and assertional information.
\end{problem}

\subsection{Problems addressed by this work}\label{sec:probaddress}

Given the technical experience outlined in Section~\ref{sec:history},
I first addressed Problem~\ref{prob:lapis}, by suggesting a simple
vocabulary for the server to tell the client what primitive operations
it is allowed to perform on a given resource. See
Section~\ref{sec:conlapis} for more on this paper.

My next concern was the implications of the
problems~\ref{prob:zipf}~and~\ref{prob:uniform} in a federated SPARQL
engine. While studying the evaluations done in the relevant literature
in the field, I decided to start an effort to address
Problem~\ref{prob:badstats} first by adopting methodology from the
statistical field of ``Design of Experiments'' (DoE). See
Section~\ref{sec:condoe} for details on this work.

Then, as \cite{buil2013sparql} confirmed my suspicions that there were
severe stability problems with SPARQL endpoints,
Problem~\ref{prob:syntacticcache} became my main focus, but clearly,
the understanding of the actual usage of caching was inadequate, so I
decided to solve Problem~\ref{prob:sanity} first. This work is
discussed in Section~\ref{sec:consanity}.

To further address Problem~\ref{prob:syntacticcache}, I decided to
write a query planner that could reuse previous results or prefetch
data the system deemed relevant, but then we realised that the current
programming paradigm was quite unhelpful when customising a query
planner for such a purpose. Rather than writing a prototypical query
engine for illustrative purposes, we wanted to integrate it into a
fully compliant SPARQL query engine, and address
problems~\ref{prob:breakdown},~\ref{prob:complexapi}~and~\ref{prob:dontjustpass}. This
is explained in Section~\ref{sec:conpush}
